### 2020-07-22 paper/coloringAppendix/algIV.tex
First draft of paper skeleton

\begin{algorithm}
	\SetKwFunction{GoodPartitionEvenH}{\textsc{Good-Partition-From-Even-Hyperprism}}
	\Indm\nonl\GoodPartitionEvenH{$G, H, M$}\\
	\KwData{$G$ -- square-free, Berge graph containing no \LGBSK \newline
		$H = (A_1, \ldots, B_3)$  -- maximal even hyperprism in $G$ \newline
		$M$ -- set of major neighbors of $H$}
	\KwResult{A good partition of $G$}
	\Indp	
	$Z \gets \bigcup\{V(C)$ : $C$ is a component of $G \setminus \{V(H) \cup M \}$ with no attachments in $H \}$  \;

	relabel strips of $H$, so that $M \cup A_1$ and $M \cup B_1$ are cliques \;

	$F_1 \gets \bigcup\{ V(C)$ : $C$ is a component of $G \setminus \{H \cup M \cup Z \}$ that
	attaches to $A_1 \cup B_1 \cup C_1$  $\}$ \;

	% $F_A \gets \bigcup \{ V(C) $ : $C$ is a component of $G \setminus \{ H \cup M \cup Z \cup F_1 \cup F_2 \cup F_3 \}$
	% that attaches to $A_1 \cup A_2 \cup A_3$  $\}$ \;

	% $F_B \gets \bigcup \{ V(C) $ : $C$ is a component of $G \setminus \{ H \cup M \cup Z \cup F_1 \cup F_2 \cup F_3 \}$
	% that attaches to $B_1 \cup B_2 \cup B_3$  $\}$ \;

	% \Assert{$F_1, F_2, F_3, F_A, F_B$ are pairwise anticomplete}
	% \Assert{$V(G) = H \cup M \cup Z \cup F_1 \cup F_2 \cup F_3 \cup F_a \cup F_B$}
	\Assert{$M$ is a clique}
	\Assert{$M \cup A_i$ is a clique for at least two values of $i$}
	\Assert{$M \cup B_j$ is a clique for at least two values of $j$}

	$K_1 \gets A_1$, $K_2 \gets M$, $K_3 \gets B_1$ \;
	$R \gets C_1 \cup F_1 \cup Z $ \;
	$L \gets G \setminus \{K_1 \cup K_2 \cup K_3 \cup R\}$ \;
	\Return $(K_1, K_2, K_3, L, R)$\;
\end{algorithm}