### 2020-07-22 paper/coloringAppendix/algVIa.tex

\subsubsection{Notes}
\begin{itemize}
	\item $T_{uv} := N_u \cap N_v$
	\item Strip $S_{uv}$ is \emph{rich} iff $S_{uv} \setminus T_{uv} \neq \emptyset$
\end{itemize}

\begin{algorithm}
	\SetKwFunction{GoodPartitionJStrip}{\textsc{Good-Partition-From-J-Strip-System}}
	\Indm\nonl\GoodPartitionJStrip{$G, J, (S, N), M$}\\
	\KwData{$G$ -- square-free, Berge graph \newline
		$J$ -- a maximal $3$-connected graph with appearance in $G$ \newline
		$(S, N)$ -- a maximal $J$-strip system \newline
		$M$ -- a set of major vertices w.r.t. $(S, N)$
	}
	\KwResult{A good partition of $G$}
	\Indp	
	$S^*_{uv} \gets S_{uv} \cup$ $($ components of $G\setminus V(S,N)$ that attach in $S_{uv}$ only $)$\;
	$T_{uv} \gets N_u \cap N_v$\;
	\Assert{$T_{uv} \setComplete N_u \setminus N_{uv}$, $T_{uv} \setComplete N_v \setminus N_{vu}$}
	\Assert{$M \cup T_{uv}$ is a clique}
	\If{$\exists S_{uv}$ -- a rich strip in $(S,N)$} {
		\If{$\exists S_{uv}$ -- a rich strip in $(S, N)$, such that
			$M \cup(N_u \setminus N_{uv})$ and $M \cup(N_v \setminus N_{vu})$ are cliques
			}{
			$K_1 \gets N_u \setminus N_{uv}$, $K_2 \gets M \cup T_{uv}$, $K_3 \gets N_v \setminus N_{vu}$\;
			$L \gets (S^*_{uv} \setminus T_{uv}) \cup ($ components of $G\setminus V(S, N)$ that attach only to $N_u$
			and those that attach only to $N_v$ $)$\;
			$R \gets G \setminus (K_1 \cup K_2 \cup K_3 \cup L)$\;
			\Return $(K_1, K_2, K_3, L, R)$\;
		}
		\Else{
			$S_{uv} \gets$ a rich strip in $(S, N)$, such that
			$M \cup(N_u \setminus N_{uv})$ is not a clique and $M \cup(N_v \setminus N_{vu})$ is a clique \;
			$K_1 \gets N_{uv} \setminus T_{uv}$, $K_2 \gets M \cup T_{uv}$, $K_3 \gets N_v \setminus N_{vu}$\;
			$R \gets (S^*_{uv} \setminus N_u) \cup ($ components of $G\setminus V(S,U)$ that attach only to $N_v$ $)$\;
			$L \gets G \setminus (K_1 \cup K_2 \cup K_3 \cup R)$\;
			\Return $(K_1, K_2, K_3, L, R)$\;
		}
	}
	\Else {
		\Assert{$\forall uv \in E(J): S_{uv} = T_{uv}, S_{uv}$ is a clique}
		$S_{uv} \gets $ any strip\;
		$K_1 \gets N_u \setminus S_{uv}$, $K_2 \gets M$, $K_3 \gets N_v \setminus S_{uv}$\;
    $L \gets S^*_{uv} \cup ($ components of $G\setminus V(S,N)$ that attach only to $N_u$ and only to $N_v$ $)$\;
		$R \gets G \setminus (K_1 \cup K_2 \cup K_3 \cup L)$\;
		\Return $(K_1, K_2, K_3, L, R)$\;
	}
\end{algorithm}