### 2020-09-07 paper/coloringAppendix/algI.tex
Algos and KT fixes

\begin{algorithm}
  \SetKwFunction{colorGraph}{\textsc{Color-Graph}}
  \Indm\nonl\colorGraph($G$)\\
  \KwData{$G$ -- square-free Berge graph}
  \KwResult{A $\chi(G)$-coloring of $G$}
  \Indp
  TODO \;
  TODO 2\;
  \Return Todo\;
\end{algorithm}