### 2020-09-07 paper/main.tex

\documentclass{report}
\usepackage{import}

\import{./}{preamble.tex}

% Language and typesetting notes:
%
% - Color, not Colour (American English)

\author{Adrian Siwiec}
\date{\today{}}
\begin{document}

\include{titlepage}
\pagebreak

\begin{abstract}
	TODOa
\end{abstract}

\listoftheorems[ignoreall,show={defn}]
\tableofcontents

\pagebreak

\chapter*{Definitions}
\input{definitions.tex}

\chapter{Perfect Graphs}
\input{perfectGraphs.tex}
\input{coloringBergeGraphs.tex}

\chapter{Implementation}
\input{implementation.tex}

\appendix
\appendixpage
\addappheadtotoc

\chapter{Perfect Graph Coloring algorithm}
\label{ch:coloringSquareFree}
\section{Notes}

In appendix we use a slightly different notation than before, with many concepts represented by inline symbols. This is intended to reduce the length of algorithm's text and simplify its analysis.

\begin{itemize}
	\item $V(X)$ -- vertices of structure $X$. Will be written as $X$ when obvious.
	\item $a - b$, when $a$ and $b$ are nodes -- $a$ and $b$ are neighbors.
	\item $a \cdots b$, when $a$ and $b$ are nodes -- $a$ and $b$ are not neighbors.
	\item $a - X$, when $a$ is a node and $X$ is a set of nodes -- $a$ has a neighbor in $X$.
	\item $a \cdots X$, when $a$ is a node and $X$ is a set of nodes -- $a$ has a nonneighbor in $X$.
	\item $a \blacktriangleleft  X$, when $a$ is a node and $X$ is a set of nodes -- $a$ is complete to $X$.
	\item $a \ntriangleleft X$, when $a$ is a node and $X$ is a set of nodes -- $a$ is anticomplete to $X$.
	\item $X \setComplete Y$, when $X$ and $Y$ are set of nodes -- $X$ is complete to $Y$.
	\item $X \setAntiComplete Y$, when $X$ and $Y$ are set of nodes -- $X$ is anticomplete to $Y$.
	\item $[n]$  -- $\{1, \ldots, n\}$.
	\item \LGBSK -- a line-graph of a biparite subdivision of $K_4$.
	\item $a \gets b$ -- let a be equal b.
	\item $a :\in X$ -- let a be equal to any element of $X$
	\item $a \xor b$ -- $a$ xor $b$

	\item Strip $S_{uv}$ is \emph{rich} iff $S_{uv} \setminus T_{uv} \neq \emptyset$
\end{itemize}
\TODO{Introduction on unknown structures - J-strip etc etc}

\section{Algorithms}

\graphAppendix{Color Graph}
\input{coloringAppendix/algI.tex}
\clearpage

\graphAppendix{Color Good Partition}
\input{coloringAppendix/algII.tex}
\clearpage

\graphAppendix{Grow Hyperprism}
\input{coloringAppendix/algIII.tex}
\clearpage

\graphAppendix{Good Partition from an even hyperprism}
\input{coloringAppendix/algIV.tex}
\clearpage

\graphAppendix{Good Partition from odd hyperprism}
\input{coloringAppendix/algV.tex}
\clearpage

\graphAppendix{Good Partition from $J$-strip system}
\input{coloringAppendix/algVIa.tex}
\clearpage

\graphAppendix{Good Partition from a special $K_4$ strip system}
\input{coloringAppendix/algVIb.tex}
\clearpage

\graphAppendix{Find a special $K_4$ system}
\input{coloringAppendix/algVII.tex}
\clearpage

\graphAppendix{Growing a $J$-strip}
\input{coloringAppendix/algVIII.tex}
\clearpage

% \graphAppendix{LAST}
% \input{coloringAppendix/algIXa.tex}
% \clearpage

% \bibliographystyle{alpha}
\printbibliography

\begin{figure}
	\input{tikzpictures/3dtest.tex}
	\caption{3d test}
\end{figure}

\end{document}