### 2020-09-13 paper/coloringAppendix/algI.tex

\begin{mynameforalgorithm}
  % \myalgorithmcommand
  \SetKwFunction{colorGraph}{\textsc{Color-Graph}}
  \Indm\nonl\colorGraph($G$)\\
  \KwData{$G$ -- square-free Berge graph}
  \KwResult{A $\omega(G)$-coloring of $G$}
  \Indp
  TODO \;
  TODO \;
  \Return Todo\mysemisemi
\end{mynameforalgorithm}