### 2020-09-14 paper/main.tex

\documentclass{report}
\usepackage{import}

\import{./}{preamble.tex}

% Language and typesetting notes:
%
% - Color, not Colour (American English)

\author{Adrian Siwiec}
\date{\today{}}
\begin{document}

\include{titlepage}
\pagebreak

\begin{abstract}
	TODOa
\end{abstract}

\listoftheorems[ignoreall,show={defn}]
\renewcommand{\listtheoremname}{List of algorithms}
\listoftheorems[ignoreall,show={alg}]
\tableofcontents

\pagebreak

\chapter*{Definitions}
\input{definitions.tex}

\chapter{Perfect Graphs}
\input{perfectGraphs.tex}

\chapter{Coloring Perfect Graphs}
\input{coloringBergeGraphs.tex}

\chapter{Implementation}
\input{implementation.tex}
\input{experiments.tex}

\appendix
\appendixpage
\addappheadtotoc

\input{appendA.tex}

% \bibliographystyle{alpha}
\printbibliography

% \begin{figure}
% 	\input{tikzpictures/3dtest.tex}
% 	\caption{3d test}
% \end{figure}

\end{document}