### 2020-09-16 paper/main.tex

\documentclass{report}
\usepackage{import}

\import{./}{preamble.tex}

% Language and typesetting notes:
%
% - Color, not Colour (American English)

\author{Adrian Siwiec}
\date{\today{}}
\begin{document}

\include{titlepage}
\pagebreak

\begin{abstract}
	Perfect graphs are a subject of intense study since their introduction in 1961. Only in 2001 the famous Perfect graph conjecture was proven to be true and in 2005 a polynomial method of determining if a graph is perfect was found. Since 1988 an algorithm is known for coloring perfect graphs, but it uses an ellipsoid method which is said to be complicated and impractical. As recently as in 2018 a polynomial algorithm for coloring perfect graphs without squares was published, that uses a combinatorial approach.

	We begin with basic definitions in Chapter 1. In Chapter 2 we introduce perfect graphs and Berge graphs, give an overview of the strong perfect graph theorem and talk about an algorithm for polynomial perfect graph recognition. Chapter 3 is a study of the ellipsoid method of coloring perfect graphs. In Chapter 4 we present our implementation of algorithms from Chapters 2 and 3, along with notes on optimization and parallelisation. Appendix A is a overview of the recent algorithm for coloring square-free perfect graphs.
\end{abstract}

\listoftheorems[ignoreall,show={defn}]
\renewcommand{\listtheoremname}{List of algorithms}
\listoftheorems[ignoreall,show={alg}]
\tableofcontents

\pagebreak

\chapter{Basic Graph Definitions}
\input{definitions.tex}

\chapter{Perfect Graphs}
\input{perfectGraphs.tex}

\chapter{Coloring Perfect Graphs}
\input{coloringBergeGraphs.tex}

\chapter{Implementation}
\input{implementation.tex}
\input{experiments.tex}

\appendix
\appendixpage
\addappheadtotoc

\input{appendA.tex}

% \bibliographystyle{alpha}
\printbibliography

% \begin{figure}
% 	\input{tikzpictures/3dtest.tex}
% 	\caption{3d test}
% \end{figure}

\end{document}