### 2020-09-16 paper/perfectGraphs.tex

Given a graph $G$, let us consider a problem of coloring it using as few colors as possible. If $G$ contains a clique $K$ as a subgraph, we must use at least $|V(K)|$ colors to color it. This gives us a lower bound for a chromatic number $\chi(G)$ -- it is always greater or equal to the cardinality of the largest clique $\omega(G)$. The reverse is not always true, in fact we can construct a graph with no triangles and requiring arbitrarily large numbers of colors (e.g. construction by Mycielski \cite{Mycielski1955}).

Do graphs that admit coloring using only $\omega(G)$ color are "simpler" to further analyze? Not necessarily so. Given a graph $G = (V, E)$, let us construct a graph $G'$ as the union of $G$ and a clique $K_{|V|}$. We can see that indeed $\chi(G') = \omega(G') = |V|$, but it gives us no indication of the structure of $G$ or $G'$. This leads us to the hereditary definition of \emph{perfect graphs}.

\begin{defn}[perfect graph]
	\label{def:perfectGraph}
	A graph $G$ is \emph{perfect} if and only if for its every induced subgraph $H$ we have $\chi(H) = \omega(H)$.
\end{defn}

The notion of perfect graphs was first introduced by Berge in 1961 \cite{CB61} and it indeed captures some of the idea of graph being ''simple'' -- in all perfect graphs the coloring problem, maximum (weighted) clique problem, and maximum (weighted) independent set problem can be solved in polynomial time \cite{grotschel1993}. Other classical NP-complete problems remain hard in perfect graphs e.g. Hamiltonian path \cite{Mller1996}, maximum cut problem \cite{Bodlaender1994} or dominating set problem \cite{Dewdney81}. There are many well-known subclasses of perfect graphs, we take a look and benchmark our implementations on some of them in \Cref{sec:experiments}

The most fundamental problem -- the problem of recognizing perfect graphs -- was open since its posing in 1961 until recently. Its solution, a polynomial algorithm recognizing perfect graphs, is a union of the strong perfect graph theorem (\Cref{sec:SPGT}) stating that a graph is perfect if and only if it is Berge and an algorithm for recognizing Berge graphs in polynomial time (\Cref{sec:recognizingBerge}).

\begin{defn}[Berge graph]
	\label{def:bergeGraph}
	A graph $G$ is \emph{Berge} if and only if both $G$ and $\overline{G}$ have no odd hole as an induced subgraph.
\end{defn}

Perfect graphs are interesting not only because of their theoretical properties, but they are also used in other areas of study e.g. integrality of polyhedra \cite{Chvtal1975, Chudnovsky2003}, radio channel assignment problem \cite{McDiarmid99, McDiarmid2000} and appear in the theory of Shannon capacity of a graph \cite{Lovasz1979}. Also, as pointed out in \cite{alfonsinPerfect2001, Chudnovsky2003}, algorithms to solve semidefinite programs grew out of the theory of perfect graphs. We will take a look at semidefinite programs and at perfect graph's relation to the Shannon capacity in \Cref{sec:ShannonCapacity}.

\section{Strong perfect graph theorem}
\label{sec:SPGT}

The first step to solve the problem of recognizing perfect graphs was the \emph{(weak) perfect graph theorem} first conjured by Berge in 1961~\cite{CB61} and then proven by Lovász in 1972~\cite{LL72}.

\begin{theorem}[Perfect graph theorem \cite{LL72}]
	\label{thm:pgt}
	A graph is perfect if and only if its complement graph is also perfect.
\end{theorem}

This theorem is a consequence of a stronger result proven by Lovász:
\begin{theorem}[\cite{LL72}, Claim 1 of \cite{GC03}]
	\label{thm:omegaalpha}
	A graph $G$ is perfect if and only if for every induced subgraph $H$, the number of vertices of $H$ is at most $\alpha(H)\omega(H)$.
\end{theorem}
% \begin{proof}
% 	First assume that $G$ is perfect -- for every induced subgraph $H$, $\omega(G) = \chi(H)$. Since the number of vertices of $H$ is at most $\alpha(H)\chi(H)$, the claim follows.

% 	The proof of the converse we show is due to Gasparyan \cite{Gasparian1996}. Assume that $G$ is not perfect. Let $H$ be a minimally imperfect subgraph of $G$ and let $n$ be the number of vertices of $H$. Let $\alpha = \alpha(H), \omega = \omega(H)$.

% 	Then, because $H$ is \emph{minimally} imperfect, it satisfies:
% 	\begin{itemize}
% 		\item $\omega = \chi(H \setminus v)$, for every $v \in H(V)$,
% 		\item $\omega = \omega(H \setminus S)$, for every stable set $S \subseteq V(H)$. 
% 	\end{itemize}

% 	Let $A_0$ be a maximal stable set of $H$. Then $|A_0| = \alpha$ and for each $s \in A_0$ fix an $\omega$-coloring of the graph $H \setminus s$.
% \end{proof}

Then, since $\alpha(H) = \omega(\overline{H})$ and $\omega(H) = \alpha(\overline{H})$ \Cref{thm:omegaalpha} implies \Cref{thm:pgt}. We skip the proof of \Cref{thm:omegaalpha}, as it is quite technical.

Odd holes are not perfect, since their chromatic number is 3 and their largest cliques are of size 2. It is also easy to see, that an odd antihole of size $n$ has a chromatic number of $\frac{n+1}{2}$ and largest cliques of size $\frac{n-1}{2}$. A graph with no odd hole and no odd antihole is called \emph{Berge} (\Cref{def:bergeGraph}) after Claude Berge who studied perfect graphs.

In 1961 Berge conjured that a graph is perfect if and only if it contains no odd hole and no odd antihole in what has become known as a strong perfect graph conjecture. In 2001 Chudnovsky et al. have proven it and published the proof in an over 150 pages long paper \citetitle{MC06} \cite{MC06}.

\begin{theorem}[Strong perfect graph theorem, Theorem 1.1 of \cite{MC06}]
	\label{thm:spgt}
	A graph is perfect if and only if it is Berge.
\end{theorem}

The proof is long and complicated. Moreover, it has little noticeable connection to the algorithm of recognizing Berge graphs we discuss later. Therefore we will discuss it very briefly following the overview by Cornuéjols \cite{GC03}.

\subsubsection{Basic classes of perfect graphs}
Bipartite graphs are perfect, since we can color them with two colors. From the theorem of König we get that line graphs of bipartite graphs are also perfect \cite{Knig1916, GC03}. From the perfect graph theorem (\Cref{thm:pgt}) it follows that complements of bipartite graphs and complement of line graphs of bipartite graphs are also perfect. We will call these four classes \emph{basic}.

\subsubsection{2-join, Homogeneous Pair and Skew Partition}
A graph $G$ has a \emph{2-join} if its vertices can be partitioned into sets $V_1$, $V_2$, each of size at least three, and there are nonempty disjoint subsets $A_1, B_1 \subseteq V_1$ and $A_2, B_2 \subseteq V_2$, such that all vertices of $A_1$ are adjacent to all vertices of $A_2$, all vertices of $B_1$ are adjacent to all vertices of $B_2$, and these are the only edges between $V_1$ and $V_2$. When a graph $G$ has a 2-join, it can be decomposed onto two smaller graphs $G_1$, $G_2$, so that $G$ is perfect if and only if $G_1$ and $G_2$ are perfect \cite{Cornujols1985}.

A graph $G$ has a \emph{homogeneous pair} if $V(G)$ can be partitioned into subsets $A_1$, $A_2$, $B$, such that $|A_1|+|A_2| \geq 3$, $|B| \geq 2$ and if a vertex $v \in B$ is adjacent to a vertex from $A_i$, then it is adjacent to all vertices from $A_i$. Chvátal and Sbihi proved that no minimally imperfect graph has a homogeneous pair \cite{Chvtal1987}.

A graph $G$ has a \emph{skew partition} if $V(G)$ can be partitioned into nonempty subsets $A, B, C, D$ such that there are all possible edges between $A$ and $B$ and no edges between $C$ and $D$. Chudnovsky et at. proved that no minimally imperfect graph has a skew partition.

The proof of \Cref{thm:spgt} is a consequence of the \emph{Decomposition theorem}:
\begin{theorem}[Decomposition theorem, stated as Theorem 4.2 of \cite{GC03}]
	\label{thm:decomposition}
	Every Berge graph $G$ is basic or has a skew partition or a homogeneous pair, or either $G$ or $\overline{G}$ has a 2-join.
\end{theorem}

See \cite{MC06} for the full proof of \Cref{thm:spgt} and \cite{GC03} for its shorter overview.

\section{Recognizing berge graphs}
\label{sec:recognizingBerge}

The following is based on the paper by \citeauthor{MC05} \citetitle*[]{MC05} \cite{MC05}. We will not provide full proof of its correctness, but we will aim to show the intuition behind the algorithm.

Berge graph recognition algorithm (later called CCLSV from the names of its authors) could be divided into two parts: first we check if either $G$ or $\overline{G}$ contain any of a number of simple forbidden structures (\Cref{SimpleStructures}). If they do, we output that graph is not Berge and stop. Else, we check if there is a near-cleaner for a shortest odd hole or for shortest odd antihole (\Cref{AmenableHoles}).

\subsection{Simple forbidden structures}
\label{SimpleStructures}

\subsubsection{Pyramids}

A \emph{pyramid} in G is an induced subgraph formed by the union of a triangle \footnote{A triangle is a clique $K_3$.} $\{b_1,b_2,b_3\}$, three paths $\{P_1, P_2, P_3\}$ and another vertex $a$, so that:

\begin{minipage}{0.60\textwidth}
\begin{itemize}
	\item $\forall_{1\leq i \leq 3}$ $P_i$ is a path between $a$ and $b_i$,
	\item $\forall_{1\leq i < j \leq 3}$ $a$ is the only vertex in both $P_i$ and $P_j$ and $b_ib_j$ is the only edge between $V(P_i)\setminus\{a\}$ and $V(P_j)\setminus\{a\}$,
	\item $a$ is adjacent to at most one of $\{b_1, b_2, b_3\}$.
\end{itemize}
\end{minipage}%
% \begin{minipage}
\begin{wrapfigure}{r}{0.32\textwidth}
	\vspace{-2.5cm}
	\input{tikzpictures/pyramid.tex}
	\caption{An example  of a pyramid.}%
	\vspace{-1.2cm}
\end{wrapfigure}
% \end{minipage}

We will say that $a$ is \emph{linked onto} the triangle $\{b_1, b_2, b_3\}$ \emph{via} the paths $P_1$, $P_2$, $P_3$. Let us notice, that a pyramid is uniquely determined by its paths $P_1$, $P_2$, $P_3$.

It is easy to see that every graph containing a pyramid contains an odd hole -- at least two of the paths $P_1$, $P_2$, $P_3$ will have the same parity.

\paragraph{Finding Pyramids.}

\begin{alg}[Test if $G$ contains a Pyramid]
	\label{alg:testPyramid}
	Input: A graph $G$.

	\noindent Output: Returns whether $G$ contains a pyramid as an induced subgraph.
\end{alg}
\begin{algorithmic}[1]
	\mProcedure{Contains-Pyramid}{$G$}
	\mForEach{triangle $b_1, b_2, b_3$} \label{line:pyramidTriangle}
		\mForEach{$s_1, s_2, s_3$, such that for $1 \leq i < j \leq 3$, $\{b_i, s_i\}$ is disjoint
			\lsx  from $\{b_j, s_j\}$ and $b_ib_j$ is the only edge between them} \label{line:pyramidTriple}
			\mIf{there is a vertex $a$, adjacent to all of $s_1, s_2, s_3$, and to at most
				\lsx one of $b_1, b_2, b_3$, such that if $a$ is adjacent to $b_i$, then $b_i = s_i$ \lsx}\label{line:a}
				\ls $M \gets V(G) \setminus \{b_1, b_2, b_3, s_1, s_2, s_3\}$
				\mForEach{$m \in M$} \label{line:pyramidSStart}
					\ls $S_1(m) \gets$ the shortest path between $s_1$ and $m$ such that
					\lsx $s_2, s_3, b_2, b_3$ have no neighbors in its interior, if such a 
					\lsx path exists.
					\ls calculate $S_2(m), S_3(m)$ similarly
					\ls $T_1(m) \gets$ the shortest path between $m$ and $b_1$, such that
					\lsx $s_2, s_3, b_2, b_3$ have no neighbors in its interior, if such a
					\lsx path exists
					\ls calculate $T_2(m), T_3(m)$ similarly
				\mEndFor \label{line:pyramidSEnd}

					\mIf{$s_1 = b_1$ \label{line:ShortestPathsStart}} \Comment{calculate all possible $P_1$ paths}
						\ls $P_1(b_1) \gets$ the one-vertex path with vertex $b_1$
						\ls \algorithmicforeach ~$m \in M$ \algorithmicdo~ $P_1(m) \gets$ \textsc{undefined}
					\mElse
						\ls $P_1(b_1) \gets $ \textsc{undefined}
						\mForEach{$m \in M$}
							\mIf{ $m$ is nonadjacent to all of $b_2, b_3, s_2, s_3$ \AND
								\lsx $S_1(m)$ and $T_1(m)$ both exist \AND
								\lsx $V(S_1(m) \cap T_1(m)) = \{m\}$ \AND
								\lsx there are no edges between $V(S_1(m) \setminus m)$ 
								\lsx and $V(T_1(m) \setminus m)$ 
								\lsx
							}
								\ls $P_1(m) \gets s_1-S_1(m)-m-T_1(m)-b_1$
							\mElse
								\ls $P_1(m) \gets$ \textsc{undefined}
							\mEndIf
						\mEndFor
					\mEndIf

				\ls assign $P_2$ and $P_3$ in a similar manner \label{line:ShortestPathsEnd}
				\ls $good\_pairs_{1,2} \gets \emptyset$ \Comment{see below for definition}
				\mForEach{$m_1 \in M \cup \{b_1\}$} \Comment{calculate good $(1, 2)$-pairs}
					\mIf{$P_1(m_1) \neq \UND$}
						\ls color black the vertices of $M$ that either belong to \label{line:pyramidColor}
						\lsx $P_1(m_1)$ or have a neighbor in $P_1(m_1)$
						\lsx color all other vertices white.
						\mForEach{$m_2 \in M \cup \{b_2\}$}
							\mIf{$P_2(m_2)$ exists and contains no black vertices \lsx} \label{line:pyramidColor2}
								\ls add $(m_1, m_2)$ to $good\_pairs_{1,2}$
							\mEndIf
						\mEndFor
					\mEndIf
				\mEndFor
				\ls calculate $good\_pairs_{1,3}$ and $good\_pairs_{2,3}$ in similar way
				\mForEach{triple $m_1, m_2, m_3$ such that $m_i \in M \cup \{b_i\}$} \label{line:m1m2m3}
					\mIf{$\forall_{1\leq i < j \leq 3}$: $(m_i, m_j)$ is a good $(i, j)$-pair}
						\ls \RETURN \TRUE \label{line:pyramidTrue}
					\mEndIf
				\mEndFor
			\mEndIf \label{line:pyramidEnd}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

With definitions as above, for $1 \leq i < j \leq 3$, we say that $(m_i, m_j)$ is a \emph{good $(i, j)$-pair}, if and only if $m_i \in M \cup \{b_i\}$, $m_j \in M \cup \{b_j\}$, $P_i(m_i)$ and $P_j(m_j)$ both exist, and the sets $V(P_i(m_i))$,$V(P_j(m_j))$ are both disjoint and $b_ib_j$ is the only edge between them. In line \ref{line:pyramidColor} we color vertices of $P_1(m_1)$ black, so that for each $m_2$ we can check if paths $P_1(m_1)$ and $P_2(m_2)$ are disjoint in $O(|V|)$ time.

It is easy to see, that if \textsc{Contains-Pyramid($G$)} outputs that $G$ contains a pyramid, it indeed does -- when we return in line \ref{line:pyramidTrue} the vertex $a$ found in line \ref{line:a} can be linked into a triangle $b_1, b_2, b_3$ via paths $P_1(m_1), P_2(m_2), P_3(m_3)$ for $m_1$, $m_2$, $m_3$ from line \ref{line:m1m2m3}. The proof of the converse is rather technical and we refer to Theorem 2.2 of \cite{MC05} for it.

Now we will proove the time complexity.

\begin{theorem}[part of Theorem 2.2 of \cite{MC05}]
	\label{thm:containsPyramid9}
	Procedure \textsc{Contains-Pyramid($G$)} works in $O(|V|^9)$ time.
\end{theorem}
\begin{proof}
	There are $O(|V|^3)$ triangles (line \ref{line:pyramidTriangle}) and $O(|V|^3)$ triples $s_1$, $s_2$, $s_3$ (line \ref{line:pyramidTriple}), so lines \ref{line:a}-\ref{line:pyramidEnd} are executed at most $O(|V|^6)$ times.

	Checking if there exists an appropriate $a$ takes linear time (line \ref{line:a}). Calculating paths $S_i$ and $T_i$ (lines \ref{line:pyramidSStart}-\ref{line:pyramidSEnd}) takes $O(|V|^2)$ time for each $m \in M$ and $O(|V|^3)$ in total. Similarly, it takes $O(|V|^3)$ time to calculate all $P_i$ paths (lines \ref{line:ShortestPathsStart}-\ref{line:ShortestPathsEnd}).

	Then, for each $m_1$ we do at most $O(|V|)$ work in line \ref{line:pyramidColor} and for each $(m_1, m_2)$ we do at most $O(|V|)$ work in line \ref{line:pyramidColor2}.

	Finally, there are $O(|V|^3)$ pairs $m_1, m_2, m_3$ and checking each takes $O(1)$ time which gives us the overall running time of $O(|V|^9)$.
\end{proof}

\subsubsection{Jewels}

% \begin{wrapfigure}[7]{r}{0.35\textwidth}
\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/jewel.tex}
	\caption{An example of a jewel.}
	\vspace{-1.8cm}
\end{wrapfigure}

Five vertices $v_1, \ldots, v_5$ and a path $P$ form a \emph{jewel} if and only if:

\begin{itemize}
	\item $v_1, \ldots, v_5$ are distinct vertices,
	\item $v_1v_2, v_2v_3, v_3v_4, v_4v_5, v_5v_1$ are edges,
	\item $v_1v_3, v_2v_4, v_1,v_4$ are nonedges,
	\item $P$ is a path between $v_1$ and $v_4$, such that $v_2, v_3, v_5$ have no neighbors in its inside.
\end{itemize}

Notice that given a jewel $v_1, \ldots, v_5, P$, if the path $P$ is odd, then $v_1$-$P$-$v_4$-$v_5$-$v_1$ is an odd hole. If the path $P$ is even, then $v_1$-$P$-$v_4$-$v_3$-$v_2$-$v_1$ is an odd hole.

Most obvious way to find a jewel is to enumerate all (possibly chordal) cycles of length 5 as $v_1, \ldots v_5$, check if it has all required nonedges and if it does, try to find a path $P$ as required. This gives us a time of $O(|V|^7)$. We could speed it up to $O(|V|^6)$ with more careful algorithm, but since whole Berge recognition algorithm takes time $O(|V|^9)$ and our testing showed that time it takes to test for jewels is negligible we decided against it.

\subsubsection{Configurations of type $\T_1$}

A configuration of type $\T_1$ is a hole of length 5. To find it, we simply iterate over all paths of length 4 and check if there exists a fifth vertex to complete the hole. See \Cref{sec:usesGeneration} for more implementation details.

\subsubsection{Configurations of type $\T_2$}

% \begin{wrapfigure}[15]{r}{0.35\textwidth}
\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/t2.tex}
	\caption{An example of a $\T_2$.}%
	\vspace{-.5cm}
\end{wrapfigure}

A configuration of type $\T_2$ is a tuple $(v_1, v_2, v_3, v_4, P, X)$, such that:
\begin{itemize}
	\item $v_1v_2v_3v_4$ is a path in $G$,
	\item $X$ is an anticomponent of the set of all $\{v_1, v_2, v_4\}$-complete vertices,
	\item $P$ is a path in $G\setminus(X \cup \{v_2, v_3\})$ between $v_1$ and $v_4$ and no vertex in $P^*$ is $X$-complete or adjacent to $v_2$ or adjacent to $v_3$.
\end{itemize}

Checking if configuration of type $\T_2$ exists in our graph is straightforward: we enumerate all paths $v_1\ldots v_4$, calculate set of all $\{v_1, v_2, v_4\}$-complete vertices and its anticomponents. Then, for each anticomponent $X$ we check if required path $P$ exists.

To prove that the existence of a configuration of type $\T_2$ implies that the graph is not Berge, we will need the following lemma:

\begin{lemma}[Roussel-Rubio Lemma \cite{RR01,MC05}]\label{lem:Roussel-Rubio}
	Let $G$ be Berge, $X$ be an anticonnected subset of $V(G)$, $P$ be an odd path $p_1\ldots p_n$ in $G\setminus X$ with length at least 3, such that $p_1$ and $p_n$ are $X$-complete and $p_2, \ldots, p_{n-1}$ are not. Then:
	\begin{itemize}
		\item $P$ is of length at least 5 and there exist nonadjacent $x, y \in X$, such that there are exactly two edges between $x, y$ and $P^*$, namely $xp_2$ and $yp_{n-1}$,
		\item or $P$ is of length 3 and there is an odd antipath joining internal vertices of $P$ with interior in $X$.
	\end{itemize}
\end{lemma}

Now, we shall prove the following:

\begin{theorem}[Theorem 6.3 of \cite{MC05}]
	If $G$ contains configuration of type $\T_2$ then $G$ is not Berge.
\end{theorem}
\begin{proof}
	Let $(v_1, v_2, v_3, v_4, P, X)$ be a configuration of type $\T_2$. Let us assume that $G$ is not Berge and consider the following:
	\begin{itemize}
		\item if $P$ is even, then $v_1, v_2, v_3, v_4, P, v_1$ is an odd hole,
		\item if $P$ is of length 3, let us name its vertices $v_1, p_2, p_3, v_4$. It follows from \Cref{lem:Roussel-Rubio}, that there exists an odd antipath between $p_2$ and $p_3$ with interior in $X$. We can complete it with $v_2p_2$ and $v_2p_3$ into an odd antihole,
		\item if $P$ is odd with the length of at least 5, it follows from \Cref{lem:Roussel-Rubio} that we have $x, y \in X$ with only two edges to $P$ being $xp_2$ and $yp_{n-1}$. This gives us an odd hole: $v_2$-$x$-$p_2$-$\ldots$-$p_{n-1}$-$y$-$v_2$.
	\end{itemize}
\end{proof}

\subsubsection{Configurations of type $\T_3$}

\begin{wrapfigure}{r}{0.35\textwidth}
	\vspace{-1cm}
	\input{tikzpictures/t3.tex}
	\caption{An example of a $\T_3$.}%
	\vspace{-2cm}
\end{wrapfigure}

A configuration of type $\T_3$ is a sequence $v_1, \ldots, v_6$, $P$, $X$, such that:
\begin{itemize}
	\item $v_1, \ldots v_6$ are distinct vertices,
	\item $v_1v_2$, $v_3v_4$, $v_1v_4$, $v_2v_3$, $v_3v_5$, $v_4v_6$ are edges, and $v_1v_3$, $v_2v_4$, $v_1v_5$, $v_2v_5$, $v_1v_6$, $v_2v_6$, $v_4v_5$ are nonedges,
	\item $X$ is an anticomponent of the set of all $\{v_1, v_2, v_5\}$-complete vertices, and $v_3$, $v_4$ are not $X$-complete,
	\item $P$ is a path of $G \setminus ( X \cup \{v_1, v_2, v_3, v_4\} )$ between $v_5$ and $v_6$ and no vertex in $P*$ is $X$-complete or adjacent to $v_1$ or adjacent to $v_2$,
	\item If $v_5v_6$ is an edge, then $v_6$ is not $X$-complete.
\end{itemize}

% The following algorithm with running time of $O(|V|^6)$ checks whether $G$ contains a configuration of type $\T_3$:

\begin{alg}[Test if $G$ contains a $\T_3$]
	\label{alg:t3}
	Input: A graph $G$.

	\noindent Output: Returns whether $G$ contains a configuration of type $\T_3$ as an induced subgraph.
\end{alg}

\begin{algorithmic}[1]
	\mProcedure{Contains-T3}{$G$}
	\mForEach{$v_1, v_2, v_5 \in V(G)$, so that $v_1v_2$ is an edge \AND
		\lsx $v_1v_5, v_2v_5$ are nonedges \label{line:t3start}}
		\ls $Y \gets$ the set of all $\{v_1, v_2, v_5\}$-complete vertices.
		\mForEach{$X$ -- an anticomponent of $Y$}
			\ls $F' \gets$ maximal connected subset containing $v_5$, such that $v_1, v_2$ \label{line:t3n4start}
			\lsx have no neighbors in $F'$ and no vertex of $F'\setminus\{v_5\}$ is $X$-complete.
			\ls $F'' \gets$ the set of all $X$-complete vertices that have a neighbor in 
			\lsx $F'$ and are nonadjacent to all of $v_1, v_2$ and $v_5$
			\ls $F \gets F' \cup F''$ \label{line:t3mid}
			\mForEach{$v_4 \in V(G) \setminus\{v_1, v_2, v_5\}$, such that $v_4$ is adjacent to $v_1$
			\lsx and not to $v_2$ and $v_5$}
				\mIf{$v_4$ has a neighbor in $F$ and a nonneighbor in $X$ \label{line:t3v4if}}
					\ls $v_6 \gets$ a neighbor of $v_4$ in $F$
					\mForEach{$v_3 \in V(G) \setminus\{v_1, v_2, v_4, v_5, v_6\}$}
						\mIf{$v_3$ is adjacent to $v_2, v_4, v_5$ and not adjacent to $v_1$ \lsx}
							\ls $P \gets$ a path from $v_6$ to $v_5$ with interior in $F'$
							\ls \RETURN \TRUE \Comment{$v_1, \ldots v_6, P, X$ is a $\T_3$} \label{line:t3found}
						\mEndIf
					\mEndFor
				\mEndIf
			\mEndFor \label{line:t3n4end}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

It is easy to see that when \Cref{alg:t3} reports a configuration of type $\T_3$ in line \ref{line:t3found}, there indeed is one. We will skip the proof that each graph containing a $\T_3$ is not Berge, as it is quite technical. See section 6.7 of \cite{MC05} for the proof.

\begin{theorem}[part of Theorem 6.4 of \cite{MC05}]
	\label{thm:t3Time}
	\Cref{alg:t3} runs in $O(|V|^6)$ time.
\end{theorem}
\begin{proof}
	There are $O(|V|^3)$ triples $(v_1, v_2, v_5)$ we enumerate in line \ref{line:t3start}. For each, there are $O(|V|)$ $X$s, so lines \ref{line:t3n4start}-\ref{line:t3n4end} are executed $O(|V|^4)$ times.

	Each execution of lines \ref{line:t3n4start}-\ref{line:t3mid} takes $O(|V|^2)$ time, and there are $O(|V|)$ $v_4$s to enumerate. Line \ref{line:t3v4if} takes $O(|V|)$ time and then there are $O(|V|)$ $v_3$ to enumerate. Each is checked in $O(1)$ time. If we find the correct one, we calculate a path $P$, but return right away, so we calculate $P$ at most once in whole algorithm.
\end{proof}

\subsection{Amenable holes.}
\label{AmenableHoles}
As we will see (\Cref{thm:amenableHoles}), if graphs $G$ and $\overline{G}$ contain no jewels, pyramids or configurations of type $\T_1, \T_2$ or $\T_3$, then every shortest odd hole in $G$ or $\overline{G}$ will have a special structure and we will call it amenable. This fact will help us determine if $G$ or $\overline{G}$ contain an odd hole. But first, let us introduce a few new definitions.

\begin{defn}[relevant triple]
	Given a graph $G$, a triple $(a, b, c)$ of vertices is called \emph{relevant} if and only if $a \neq b$ (but possibly $c \in \{a, b\}$) and $G[\{a,b,c\}]$ is an independent set.
\end{defn}

\begin{defn}[C-major vertices]
  Given a shortest odd hole $C$ in $G$, a vertex $v \in V(G) \setminus V(C)$ is $C$-major if and only if the set of its neighbors in $C$ is not contained in any 3-vertex path of $C$.
\end{defn}

\begin{defn}[clean odd hole]
  An odd hole $C$ of $G$ is \emph{clean} if and only if no vertex in $G$ is $C$-major.
\end{defn}

\begin{defn}[cleaner]
  Given a shortest odd hole $C$ in $G$, a subset $X \subseteq V(G)$ is a \emph{cleaner for $C$} if and only if $X \cap V(C) = \emptyset$ and every $C$-major vertex belongs to $X$.
\end{defn}

Let us notice, that if $X$ is a cleaner for $C$, then $C$ is a clean hole in $G \setminus X$.

\begin{defn}[near-cleaner]
  Given a shortest odd hole $C$ in $G$, a subset $X \subseteq V(G)$ is a \emph{near-cleaner for $C$} if and only if $X$ contains all $C$-major vertices and $X \cap V(C)$ is a subset of vertex set of some 3-vertex path of $C$.
\end{defn}

\begin{defn}[amenable odd hole]
  An odd hole $C$ of $G$ is \emph{amenable} if and only if it is a shortest odd hole in $G$, $|E(C)| \geq 7$, and for every anticonnected set $X$ of $C$-major vertices there is a $X$-complete edge in $C$.
\end{defn}

\begin{figure}
  \begin{minipage}{.5\textwidth}
      \input{tikzpictures/amenable.tex}
			\captionof{figure}{Amenable odd hole}
			\label{fig:amenableOddHole}
  \end{minipage}%
  \begin{minipage}{.5\textwidth}
      \input{tikzpictures/notAmenable.tex}
			\caption{Not amenable odd hole}
			\label{fig:notAmenableHole}
  \end{minipage}
  \end{figure}

Some pictures might be helpful. \Cref{fig:amenableOddHole} shows an example of an amenable odd hole $C = v_1$-$\ldots$-$v_7$-$v_1$. It is not a clean hole -- vertices $v_8$ and $v_9$ are $C$-major, and $\{v_8, v_9, v_1, v_2, v_3\}$ is an axample of a near-cleaner for $C$.
\Cref{fig:notAmenableHole} shows an odd hole $C = v_1$-$\ldots$-$v_7$-$v_1$ that is not amenable, because there is no $\{v_8, v_9\}$-complete edge in $C$.

\begin{theorem}[Theorem 8.1 of \cite{MC05}]
	\label{thm:amenableHoles}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then every shortest odd hole in $G$ is amenable.
	% \TODO{Any ideas on what else to say here? List all 9 steps?}
	% \TODO{Dobre pytanie. Te 9 kroków to da się zawrzeć w 1-2 zdaniach każde? Jeśli nie, to tylko napisałbym ogólnikowo, że "jest sekwencja kroków, która to robi, tu jest odnośnik"}
\end{theorem}

The proof of this theorem is quite technical and we will not discuss it here. See section 8 of \cite{MC05} for the proof. We can see an example of the contraposition of the \Cref{thm:amenableHoles} working in \Cref{fig:notAmenableHole} -- vertices $v_1$, $v_2$, $v_8$, $v_6$, $v_7$ and a path $v_1$-$v_9$-$v_5$-$v_6$ form a jewel.

With \Cref{thm:amenableHoles} we can describe the rest of the algorithm.

\begin{alg}[List possible near cleaners]
	\label{alg:listNearCleaners}
	Input: A graph $G$.

	\noindent Output: $O(|V|^5)$ subsets of $V(G)$, such that if $C$ is an amenable hole in $G$, then one of the subsets is a near-cleaner for $C$.
\end{alg}

\begin{algorithmic}[1]
	\mProcedure{List-Possible-Near-Cleaners}{$G$}
	\ls $\mathcal{X} \gets \emptyset$
	\mForEach{relevant triple $(a, b, c)$}
		\ls $A \gets$ anticomponents of $N(\{a, b\})$
		\mIf{$c$ is $N(\{a, b\})$-complete}
			\ls $r \gets 0$
		\mElse
			\ls $r \gets$ the cardinality of the largest set of $A$ that contains 
			\lsx a nonneibhbor of $c$ 
		\mEndIf
		\ls $Y \gets \emptyset$ 
		\mForEach{$A_i \in A$}
			\mIf{$|V(A_i)| > r$}
				\ls $Y \gets Y \cup A_i$
			\mEndIf
			\mIf{$c \in A_i$}
				\ls $W \gets A_i$
			\mEndIf
		\mEndFor
		\ls $Z \gets$ the set of all $(Y \cup W)$-complete vertices
		\ls $X(a, b, c) \gets Y \cup Z$ \Comment{we use $X(a, b, c)$ in the proof of correctness}
		\ls $\mathcal{X} \gets \mathcal{X} \cup X(a, b, c)$
	\mEndFor
	\ls $\mathcal{N} \gets \emptyset$
	\mForEach{edge $uv \in E(G)$}
		\ls $\mathcal{N} \gets \mathcal{N} \cup N(\{u, v\})$
	\mEndFor
	\ls $\mathcal{R} \gets \emptyset$
	\mForEach{$X_i \in \mathcal{X}$}
		\mForEach{$N_j \in \mathcal{N}$}
			\ls $\mathcal{R} \gets \mathcal{R} \cup (X_i \cup N_j)$ \label{line:listNCcalcR}
		\mEndFor
	\mEndFor
	\ls \RETURN $\mathcal{R}$
	\mEndProcedure
\end{algorithmic}

To prove the correctness of \Cref{alg:listNearCleaners} we will need the following theorem.

\begin{theorem}[Theorem 9.1 of \cite{MC05}]
	\label{thm:91}
	Let $C$ be a shortest odd hole in $G$, with length at least seven. Then there is a relevant triple $(a, b, c)$ of vertices such that
	\begin{itemize}
		\item the set of all $C$-major vertices not in $X(a, b, c)$ is anticonnected,
		\item $X(a, b, c) \cap V(C)$ is a subset of the vertex set of some 3-vertex path of $C$.
	\end{itemize}
\end{theorem}

\begin{theorem}[Theorem 9.2 of \cite{MC05}]
	\label{thm:92}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then the result of the \Cref{alg:listNearCleaners} on the graph $G$ is a set $\mathcal{R}$, such that for every shortest amenable hole $C$, there is a near-cleaner for $C$ in $\mathcal{R}$.
\end{theorem}
\begin{proof}
	Let us suppose that $C$ is an amenable hole in $G$. By \Cref{thm:91}, there is a relevant triple $(a, b, c)$ satisfying that theorem. Let $T$ be the set of all $C$-major vertices not in $X(a,b,c)$. From \Cref{thm:91} we get that $T$ is anticonnected. Since $C$ is amenable, there is an edge $uv$ of $C$ that is $T$-complete, and therefore $T \subseteq N(u, v)$. But then $N(u, v) \cup X(a, b, c) \in \mathcal{R}$ is a near-cleaner for $C$. Therefore the output of the \Cref{alg:listNearCleaners} is correct.
\end{proof}

\begin{alg}[Test possible near cleaner]
	\label{alg:testNearCleaner}
	Input: A graph $G$ containing no simple forbidden structure, and a subset $X \subseteq V(G)$.

	\noindent Output: Determines one of the following:
	\begin{itemize}
		\item $G$ has an odd hole,
		\item There is no shortest odd hole $C$ such that $X$ is a near-cleaner for $C$.
	\end{itemize}
\end{alg}

\begin{algorithmic}[1]
	\mProcedure{Test-Near-Cleaner}{$G, X$}
		\mForEach{pair $x, y \in V(G)$, such that $x \neq y$}
			\mIf{there is a path between $x$ and $y$ with no internal vertices in $X$ \lsx}
				\ls $R(x, y) \gets$ the shortest path $R(x, y)$ between $x, y$ with no
				\lsx internal vertex in $X$.
				\ls $r(x, y) \gets$ the length of $R(x, y)$
			\mElse
				\ls $R(x, y) \gets \textsc{undefined}$
				\ls $r(x, y) \gets \infty$
			\mEndIf
			\mForEach{$y_1 \in V(G) \setminus X$} \label{line:cudaStart}
				\mForEach{3-vertex path $x_1-x_2-x_3$ of $G\setminus y_1$} \label{line:3vertex}
					\mIf{$R(x_1, y_1), R(x_2, y_2)$ both exist} \label{line:cudaWorkStard}
						\ls $y_2 \gets$ the neighbor of $y_1$ in $R(x_2, y_1)$ \label{line:definey2}
						\mIf{$r(x_2, y_1) = r(x_1, y_1) + 1 = r(x_1, y_2)$ \AND
						\lsx $r(x_2, y_1) \leq min(r(x_3, y_1), r(x_3, y_2))$ \label{line:manyconditions}}
						\ls \RETURN an odd hole found
						\mEndIf
					\mEndIf \label{line:cudaWorkEnd}
				\mEndFor
			\mEndFor \label{line:cudaEnd}
		\mEndFor
		\ls \RETURN no odd hole found
	\mEndProcedure
\end{algorithmic}

\begin{theorem}[Theorem 5.1 of \cite{MC05}]
	\label{thm:testNCWorks}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Let $X$ be a subset of $V(G)$. If \Cref{alg:testNearCleaner} reports an odd hole, then $G$ contains an odd hole, else there is no shortest odd hole $C$ in $G$, such that $X$ is a near-cleaner for $C$.
\end{theorem}
\begin{proof}
	Below we will prove that if the \Cref{alg:testNearCleaner} reports an odd hole in $G$, there indeed is one. The rest of the proof is more complicated, see section 4 and Theorem 5.1 of \cite{MC05}.

	Let us suppose that there is a choice of $x_1$, $x_2$, $x_3$, $y_1$ satisfying conditions in lines \ref{line:cudaWorkStard} and \ref{line:manyconditions}, let $y_2$ be defined as in line \ref{line:definey2} and let $k = r(x_2, y_1)$. We claim that $G$ contains an odd hole.

	{\makeatletter
	\let\par\@@par
	\par\parshape0
	\everypar{}\begin{wrapfigure}{r}{0.35\textwidth}
		\input{tikzpictures/nearCleanerFound.tex}
		\caption{An odd hole is found}%
		\vspace{-0.3cm}
		\end{wrapfigure}
		Let $R(x_1, y_1) = p_1$-$\ldots$-$p_k$, and let $R(x_2, y_1) = q_1$-$\ldots$-$q_{k+1}$, where $p_1 = x_1$, $p_k = q_{k+1} = y_1$, $q_1 = x_2$ and $q_k = y_2$. From the definition of $R(x_1, y_1)$ and $R(x_2, y_1)$, none of $p_2, \ldots, p_{k-1}, q_2, \ldots, q_k$ belong to $X$. Also, from the definition of $y_1$, $y_1 \notin X$.

		Since $r(x_1, y_1) = r(x_2, y_1) - 1$ and since $x_1$, $x_2$ are nonadjacent it follows that $x_2$ does not belong to $R(x_1, y_1)$ and $x_1$ does not belong to $R(x_2, y_1)$. Since $r(x_3, y_1), r(x_3, y_2) \geq k = r(x_1, y_2)$ it follows that $x_3$ does not belong to $R(x_1, y_1)$ or to $R(x_2, y_1)$, and has no neighbors in $R(x_1, y_1) \cup R(x_2, y_1)$ other than $x_1$, $x_2$. Since $r(x_1, y_2) = k$ we get that $y_2$ does not belong to $R(x_1, y_1)$.

		We claim first that the insides of paths $R(x_1, y_1)$ and $R(x_2, y_1)$ have no common vertices. For suppose that there are $2 \leq i \leq k-1$ and $2 \leq j \leq k$ that $p_i = q_j$. Then the subpaths of these two paths between $p_i$, $y_1$ are both subpaths of the shortest paths, and therefore have the same length equal to $j=i+1$. So $p_1$-$\ldots$-$p_1$-$q_{j+1}$-$\ldots$-$q_k$ contains a path between $x_1$, $y_2$ of length $\leq k-2$, contradicting that $r(x_1, y_2) = k$. So $R(x_1, y_1)$ and $R(x_2, y_1)$ have no common vertices except $y_1$.\par}%

	If there are no edges between $R(x_1, y_1) \setminus \{y_1\}$ and $R(x_2, y_1) \setminus \{y_1\}$ then the union of these two paths and a path $x_1$-$x_3$-$x_2$ forms an odd hole, so the answer is correct.

	Suppose that $p_iq_j$ is an edge for some $1\leq i \leq k-1$ and $1 \leq j \leq k$. We claim $i \geq j$. If $j = 1$ this is clear so let us assume $j > 1$. Then there is a path between $x_1$, $y_2$ within $\{p_1, \ldots , p_i, q_j, \ldots q_k\}$ which has length $\leq k-j+1$ and has no internal vertex in $X$ (since $j > 1$); and since $r(x_1, y_2) = k$, it follows that $k-j+i \geq k$, that is, $i \geq j$ as claimed.

	Now, since $x_1$, $x_2$ are nonadjacent, we have $i \geq 2$. But also $r(x_2, y_1) \geq k$ and so $j+k-i \geq k$, which gives us $j \geq i$. So we get $i=j$. Let us choose $i$ minimum, then $x_3$-$x_1$-$\ldots$-$p_i$-$q_i$-$\ldots$-$x_2$-$x_3$ is an odd hole, which was what we wanted.
\end{proof}

\subsection{Algorithm for perfect graph recognition}

To summarize, we present a complete algorithm for recognizing perfect graphs.
\begin{alg}[Determine if a graph is perfect]
	\label{alg:testNearCleaner}
	Input: A graph $G$.

	\noindent Output: Determines if $G$ is perfect.
	\label{alg:bergeness}
\end{alg}

\begin{algorithmic}[1]
	\mProcedure{Check-Is-Perfect}{$G$}
		\mIf{\texttt{Contains-Pyramid}$(G)$ \OR \texttt{Contains-Jewel}$(G)$ \OR \lsx 
		\texttt{Contains-T1}$(G)$ \OR \texttt{Contains-T2}$(G)$ \OR \texttt{Contains-T3}$(G)$ \label{line:simpleStart}}
			\ls \RETURN $G$ is not perfect
		\mEndIf
		\mIf{\texttt{Contains-Pyramid}$(\overline{G})$ \OR \texttt{Contains-Jewel}$(\overline{G})$ \OR \lsx 
		\texttt{Contains-T1}$(\overline{G})$ \OR \texttt{Contains-T2}$(\overline{G})$ \OR \texttt{Contains-T3}$(\overline{G})$}
			\ls \RETURN $G$ is not perfect
		\mEndIf \label{line:simpleEnd}
		\ls $\mathcal{R} \gets $\texttt{List-Possible-Near-Cleaners}$(G)$ \label{line:getR}
		\mForEach{$X \in \mathcal{R}$} \label{line:testXStart}
			\mIf{\texttt{Test-Near-Cleaner}$(G, X)$ reports an odd hole}
				\ls \RETURN $G$ is not perfect
			\mEndIf
		\mEndFor \label{line:testXEnd}
		\ls $\mathcal{R'} \gets $\texttt{List-Possible-Near-Cleaners}$(\overline{G})$
		\mForEach{$X \in \mathcal{R'}$}
			\mIf{\texttt{Test-Near-Cleaner}$(\overline{G}, X)$ reports an odd hole}
				\ls \RETURN $G$ is not perfect
			\mEndIf
		\mEndFor
		\ls \RETURN $G$ is perfect
	\mEndProcedure
\end{algorithmic}

\begin{theorem}[Theorem 10.1 of \cite{MC05}]
	Given a graph $G = (V, E)$, \Cref{alg:bergeness} works in $O(|V|^9)$ time and determines if $G$ is perfect.
\end{theorem}
\begin{proof}
	First, in lines \ref{line:simpleStart}-\ref{line:simpleEnd} we test whether $G$ or $\overline{G}$ contain a pyramid, a jewel or a configuration of type $\T_1, \T_2$ or $\T_3$. If it is not the case, then by \Cref{thm:amenableHoles} every shortest odd hole in $G$ and in $\overline{G}$ is amenable.

	By \Cref{thm:92} in line \ref{line:getR} we get a set $\mathcal{R}$, such that if $G$ contains an amenable hole $C$, one of the sets $X \in \mathcal{R}$ is a near-cleaner for $C$. In lines \ref{line:testXStart}-\ref{line:testXEnd} we test each $X \in \mathcal{R}$ and if an odd hole is reported, then by \Cref{thm:testNCWorks} $G$ is not Berge. If for all $X \in \mathcal{R}$ an odd hole is not reported, then by \Cref{thm:amenableHoles}, \Cref{thm:92} and \Cref{thm:testNCWorks} there is no odd hole in $G$. We run the same checks on $\overline{G}$ and if again we do not detect an odd hole, then $G$ is Berge and by \Cref{thm:spgt} $G$ is perfect.

	Finally, let us add up the total running time. By \Cref{thm:containsPyramid9} checking if there is a pyramid in $G$ or $\overline{G}$ takes $O(|V|^9)$ time. Testing for jewels takes $O(|V|^7)$ time, testing for a configuration of type $\T_1$ takes $O(|V|^5)$ time and testing for configuration of type $\T_2$ takes $O(|V|^6)$ time. By \Cref{thm:t3Time} testing for a configuration of type $\T_3$ takes $O(|V|^6)$ time. Running \Cref{alg:listNearCleaners} takes $O(|V|^5)$ time and returns $O(|V|^5)$ subsets. For each subset running \Cref{alg:testNearCleaner} takes $O(|V|^4)$ time, so testing all possible near-cleaners takes $O(|V|^9)$ time.

	The overall running time is $O(|V|^9)$ with testing for pyramids and testing all possible near-cleaners being parts with biggest time complexity.
\end{proof}