### 2020-09-21 paper/coloringAppendix/algI.tex
Overweekend fixes

\begin{mynameforalgorithm}
  % \myalgorithmcommand
  \SetKwFunction{colorGraph}{\textsc{Color-Graph}}
  \Indm\nonl\colorGraph($G$)\\
  \KwData{$G$ -- square-free Berge graph}
  \KwResult{A $\omega(G)$-coloring of $G$}
  \Indp
  \If(\tcp*[f]{\Cref{thm:getPrism}}){$G$ is an Artemis graph}{
    \Return coloring of an Artemis graph $G$ \tcp*{\Cref{thm:colorArtemis}}
  }
  $H \gets$ an induced prism of $G$ \tcp*{\Cref{thm:getPrism}}

  \While{$P = $ \textsc{undefined}}{

    \If{$\exists$ a component of $G \setminus(H \cup M)$ with a set of attachments in $H$ not local}{
      $F \gets$ a minimal component of $G \setminus(H \cup M)$ with a set of attachments in $H$ not local

      $M \gets \{v: v$ is a major neighbor of H$\}$\;

      $H' \gets$ \textsc{Grow-Hyperprism}$(G, H, M, F)$\;
      \If{$H'$ is  a \LGBSK}{
        $J \gets H'$\;
        $(S, N) \gets$ a $J$-strip system\;
      }\Else{
        $H \gets H'$\;
      }
    }
    \Else(\tcp*[f]{$H$ is a maximal hyperprism}){
      $M \gets \{v: v$ is a major neighbor of H$\}$\;
      \If{$H$ is an even hyperprism}{
        $P \gets$ \textsc{Good-Partition-From-Even-Hyperprism}$(G, H, M)$
        \textbf{break}
      }
      \Else{
        $P \gets$ \textsc{Goog-Partition-From-Odd-Hyperprism}$(G, H, M)$\;
        \textbf{break}
      }
    }
    \If(\tcp*[f]{a $J$-strip system was encountered}){$J \neq$ \textsc{undefined}}{
      $J', (S', N') \gets$ \textsc{Growing-J-Strip}$(G, J, (S, N))$\;
      $M \gets$ a set of major vertices w.r.t. $(S, N)$\;
      \If{$J', (S', N')$ is a special $K_4$ system}{
        $P \gets$ \textsc{Good-Partition-From-Special-Strip-System}$(G, (S, N), M)$
      }\Else{
        $P \gets$ \textsc{Good-Partition-From-J-Strip-System}$(G, J, (S, N), M)$
      }
    }
  }
  \Return \textsc{Color-Good-Partition}$(G, P)$\;
  \vspace{-5cm}
\end{mynameforalgorithm}